---
title: "loops_and_functions.Rmd"
author: "Adelaide Robinson"
date: '2022-08-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Create two sequences, one called weekdays that contains days of the week ("Monday", "Tuesday", "Wednesday", etc.) and one called transects that contains the series of transect names "Transect A", "Transect B,"Transect C". Write a nested for loop that creates a matrix containing the following: we want our output to be a vector in which we iterate across weekdays first populating rows, then move on to the next transect

```{r}
weekdays <-c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
transects <- c("Transect A", "Transect B", "Transect C")

matrix <- matrix(nrow = length(weekdays), ncol = length(transects))
for(i in seq_along(transects)){
  for(j in seq_along(weekdays)){
    pudding <- paste(weekdays[j],"-",transects[i])
    matrix[j,i]<- pudding
  }
}

```

Task 2

Write a function called force that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2 (recall: F=ma), and returns a statement "The resulting force is \_\_\_ Newtons."

```{r}
force <- function(mass,acceleration){
  paste("The resulting force is", mass*acceleration, "newtons")
  
  }
```

Task 3

Write a function called name_column that, given a data frame name and a number, will return the column name associated with that column number. Add an error message if the column number is \< 1, or if the number exceeds the number of columns in the data frame. Test the function using the mtcars data frame.

```{r}
name_col <- function(d_f, number){
  if(number < 1){
    stop("error")
      if(number > ncol(d_f)){
        stop("big error")
      }
        
    
  }
  print(colnames(d_f[number]))
  
}
```

Task 4 Recreate the table above as a data frame stored as fish_parms. Then, write a function called fish_weight that allows a user to only enter the common name (argument fish_name) and total length (argument tot_length) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and lengths.

Now, try creating a vector of lengths (e.g. 0 to 100, by increments of 1) and ensuring that your function will calculate the fish weight over a range of lengths for the given species (try this for milkfish, storing the output weights as milkfish_weights.

```{r}
fish_parms <- tribble( ~ sci_name, ~common_name, ~a_est, ~b_est,
                    "Chanos chanos", "Milkfish",0.0905, 2.52,
         "Sphyraena barracuda", "Great barracuda", 0.0181,3.27,
         "Caranx ignobilis", 	"Giant trevally", 0.0353, 3.05 
            )
fish_weight <- function(fish_name, tot_length){
  my_sub <- filter(fish_parms, common_name == fish_name)
  my_sub$a_est*tot_length^my_sub$b_est
}
milkfish_weights <- seq(from = 0, to = 100, by = 0.1)

```

wave power

```{r}
wave_power <- function(height, period){
  0.5 * height * period
}
heights <- seq(from = 0, to = 3, by = 0.2)
wave_power(heights, 8)
```

## Task 6 

```{r}
brenville_wave_power <- function(height,period, depth){
if( depth > 12){
  message("deep ocean equation used")
  0.5 * height ^2 * period
} else if (depth <= 12){
  message("Shallow ocean equation used")
  0.81 * height ^2 * period
}
}
```
